---
title: "Impact of immigration policy on voters for election"
author: "Azad Naik"
date: "24th January 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
bibliography: Reference.bib
link-citations: yes
---

<!-- For more info on RMarkdown see http://rmarkdown.rstudio.com/ -->

<!-- Enter the code required to load your data in the space below. The data will be loaded but the line of code won't show up in your write up (echo=FALSE) in order to save space-->
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r echo=FALSE, cache=FALSE, include=FALSE}
#Load your data here
library(tidyverse)
library(knitr)
library(dplyr)
library(rmarkdown)
library(gridExtra)
anes <- read_csv("data/anes.csv")
anes <- anes %>% mutate_if(is.character, factor) # strings as factors
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE) # For wrapping the R code in the pdf document
```

<!-- In the remainder of the document, add R code chunks as needed -->

### Introduction:

How much does the immigration policy issue affect the voters in choosing candidates based on the previous data results from 2008 election.

This is an important consideration because many voters account this factor in making decision to vote for the candidates @reference1.

### Data:

Data is extracted from the American National Election Studies survey. To make analysis easier raw dataset has been pre-processed that is made available from the ANES site @reference2. 

### Exploratory data analysis:

Since we are interested in finding the relationship between the immigration policy on the candidates chosen, we only focus on two columns (or variables) in the whole dataset - interest_whovote2008 and immig_policy. 

Analysis Details:

Identify list of usable rows for analysis. For data to be meaningful both column (or variables) should have meaningful data i.e. shoudn't have NA entry. 

```{r}
voterimmig <- select(anes, interest_whovote2008, immig_policy) %>% filter(!is.na(interest_whovote2008), !is.na(immig_policy))
dim(voterimmig)
```
Verify the details of the extracted variables.

```{r}
head(voterimmig)
```
Determine percentage of votes received by Obama, Mccain and others in sample.

```{r}
print(percentageVote <- table(voterimmig$interest_whovote2008))
```
Determine distribution of votes to each candidate based on different immigration policy.

```{r}
prop.table(table(voterimmig[voterimmig$immig_policy=="Allow Unauthorized Immigrants To Remain In The United States ...Without Penalties",]$interest_whovote2008))
```

```{r}
prop.table(table(voterimmig[voterimmig$immig_policy=="Allow Unauthorized Immigrants To Remain In The United States ...Certain Requirements",]$interest_whovote2008))
```

```{r}
prop.table(table(voterimmig[voterimmig$immig_policy=="Have A Guest Worker Program That Allows Unauthorized Immigrants To Remain .",]$interest_whovote2008))
```

```{r}
prop.table(table(voterimmig[voterimmig$immig_policy=="Make All Unauthorized Immigrants Felons And Send Them Back To Their Home Country.",]$interest_whovote2008))
```

From the above results its clear that the majority of the voters who had a view to allow "unauthorized immigrants to remain in the US" support Obama while others support McCain. This is because the Obama party (i.e. Democrats) supports the implication of rules which facilitates unauthorized immigrants to remain within US.

Visual representation:
```{r}
p1 <- ggplot(voterimmig, aes(x=immig_policy, fill=interest_whovote2008))+scale_x_discrete(label=abbreviate)+geom_bar()+xlab("Voters Type")+ylab("Voters Count")
p2 <- ggplot(voterimmig, aes(x=immig_policy, fill=interest_whovote2008))+scale_x_discrete(label=abbreviate)+geom_bar(position='fill')+xlab("Voters Type")+ylab("Voters Proportion")
grid.arrange(p1, p2)
```
From the above visualization its clear that "allowing unauthorized immigrants to remain in US" support Obama whereas others support Mccain.

### Inference:
Does immigration policy affects voters decision about which candidate to vote.

#### Hypothesis:
H0: Immigration Policy and voting decisions are independent.

HA: Immigration Policy and voting decisions are dependent.

#### Findings:
```{r}
print(tbl <-table(voterimmig$interest_whovote2008, voterimmig$immig_policy))
```
Independence test using Pearson's Chi-squared test.
```{r}
print( ch <- chisq.test(tbl))
```
Based on above results its 0 probability that voting decisions and immigration policy are independent.


### Conclusion:
In this assessment we checked if various immigration policy affects the voting decision for the voters. Based on our exploratory analysis and hypothesis testing we figured out that indeed immigration policy is one of the major factor that influnce voting decisions.

### References: